Perihelion advance of Mercury

### 2020-10-08 perihelion_advance_of_mercury.Rmd
typo
typos

In the above diagram we see over 15.4 orbits of Mercury, exhibiting no
visible deviation from an exactly closed orbital path.  Nevertheless,
the orbit is not closed as Mercury has a tiny amount of orbital
precession.  Note that the Schwarzschild radius would be microsopic at
this scale (the size of the central disk is dominated by the width of
the plotting line used for the event horizon).  To calculate this
precession we need to simulate the motion for a longer time:

@@

In the above plot, note the differing axis scales.  The start point is
shown as a solid circle, subseqent points shown with open circles.
Lines connect consecutive points in the simulation.  It is difficult
to see evidence for precession in this plot because the points
correspond to regular snapshots of Mercury's position in successive
orbits, and the time evolution is not shown.

To proceed we need to find the perihelion (or an approximation to it)
somehow, and show that the estimated perihelion systematically
increases with time.  Given the numerics, we find points on the
simulated trajectory which are a local minimum distance from the sun.
This is straightfoward numerically; we identify iterations $i$ (which
have radius `r[i]`) satisfying both `r[i-1]<r[i]` and `r[i]>r[i+1]`:

>>>>

In the above diagram we see over 15.4 orbits of Mercury, exhibiting no
visible deviation from an exactly closed orbital path.  Nevertheless,
the orbit is not closed as Mercury has a tiny amount of orbital
precession.  Note that the Schwarzschild radius would be microscopic at
this scale (the size of the central disk is dominated by the width of
the plotting line used for the event horizon).  To calculate this
precession we need to simulate the motion for a longer time:

@@

In the above plot, note the differing axis scales.  The start point is
shown as a solid circle, subsequent points shown with open circles.
Lines connect consecutive points in the simulation.  It is difficult
to see evidence for precession in this plot because the points
correspond to regular snapshots of Mercury's position in successive
orbits, and the time evolution is not shown.

To proceed we need to find the perihelion (or an approximation to it)
somehow, and show that the estimated perihelion systematically
increases with time.  Given the numerics, we find points on the
simulated trajectory which are a local minimum distance from the sun.
This is straightforward numerically; we identify iterations $i$ (which
have radius `r[i]`) satisfying both `r[i-1]<r[i]` and `r[i]>r[i+1]`:
